---
title: 'Week 9 Lab: Part II Chi Square, with lizards tails'
author: "Will Geiken"
date: "11/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
library(broom)
```

## Read in data
```{r}
lizards <- read.csv(here("data","lizards.csv"))
```

Get counts of lizards by species
```{r}
lizards %>% 
  count(spp) %>% 
  arrange(-n) #arranges counts by decreasing order
```

## Make a subset

- UTST lizards
- Sites with 3 highest trap populations: IBPE (enclosure grassland), RABB (rabbit mesquite site), "EAST" (east tarbush site)
- Tail condition is recorded (W = whole, B = broken)

```{r}
utst <- lizards %>% 
  filter(spp == "UTST") %>% 
  filter(site %in% c("IBPE", "RABB", "EAST")) %>% 
  filter(tail %in% c("B", "W"))

#%in% any in that vector
# use unique(utst$site) to check that the filter worked
```

## Find counts and proportions of lizards with broken or whole tails at each site

```{r}
#normal methodology count(), gives longform table, could use pivotwider() to split it into contigency table
utst %>% 
  count(site, tail)
```

```{r}
#cool way to get it ready for Chi Square/communication
tail_counts <- utst %>% 
  janitor::tabyl(site, tail)

#now make a table with counts and proportions by row, you can specify to calculate by columns too but it requires a note
tail_proportions <- tail_counts %>% 
  adorn_percentages() %>% 
  janitor::adorn_pct_formatting(digits = 2) %>% 
  janitor::adorn_ns()

#converts to %, rounds to 2 digits, and adds actual counts
#rabb site seems to have a higher proportion of whole tails than the other sites, so we want to test if that's significant (tail breakage =/= independent of site)
```

## Convert to a contigency table so that we can run Chi-squre

Use `column_to_rowname()`.

```{r}
#makes the site variables into the row names
lizard_ct <- tail_counts %>% 
  column_to_rownames(var = "site")
```

## Chi-square
Asking: are tail breakage and site independent, are they so different that we can reject the null hyp of independence and conclude that tail breakage is dependent on site

```{r}
lizard_x2 <- chisq.test(lizard_ct)
lizard_x2

lizard_tidy <- broom::tidy(lizard_x2)

#df = degrees of freedom = (rows - 1)*(columns - 1)
# chi-sq value of 13.064 with p-value 0.001456, yes we have evidence to reject the null hyp that .. there is a significant association between site and tail breakage
```

There is a significant association (i.e. non-independence) between site and lizard tail breakage ($\chi$^2^(`r lizard_tidy$parameter`) = `r lizard_tidy$statistic`, *p* = `r lizard_tidy$p.value`).
