---
title: "Part 2 - Chi-square introduction"
output: html_document
---

## Overview

In Part 2, we'll do an introduction to chi-square test for independence in R using lizard tail breakage data recorded at Jornada Basin Long Term Ecological Research site. 

**Data from:** Lightfoot, D. 2017. Lizard pitfall trap data (LTER-II, LTER-III) ver 36. Environmental Data Initiative. https://doi.org/10.6073/pasta/ff37b4cace16a9943575f3fd7067064e (Accessed 2020-07-23).

The data are in data/lizards.csv of this project. 

## Getting started

- Add your own new .Rmd to the src/ folder for Part 2 to follow along on your own (note that the instructor key is also included in src/ - but you should be typing this out on your own for improved learning and retention!)

- Complete the following to practice the tools & concepts covered in Week 9 lectures (chi-square)

## Step 2. Attach required packages

For this part, you'll need:

- `tidyverse`
- `here`
- `janitor`
- `broom`

```{r setup, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
library(broom)
```

## Step 3. Get data & take a look

Read in the lizards.csv data and take a look (using `View(lizards)` in the Console) to familiarize yourself with the data structure: 
```{r}
lizards <- read_csv(here("data","lizards.csv"))
```

And for fun, let's find the counts of each lizard species in the data: 
```{r}
lizards %>% 
  count(spp) %>% 
  arrange(-n)
```

The most commonly observed lizard here (UTST) is *Uta stansburiana*, or commonly the side-blotched lizard. We'll continue by exploring just that lizard. 

## Make a subset

Make a subset from `lizards` that only contains observations for: 
 
- side-blotched lizards (UTST)
- at sites "IBPE" (enclosure grassland site), "RABB" (rabbit mesquite site), and "EAST" (east tarbush site)
- where the tail condition is recorded as either "W" (whole) or "B" (broken)

```{r}
utst <- lizards %>% 
  filter(spp == "UTST") %>% 
  filter(site %in% c("IBPE","RABB","EAST")) %>% 
  filter(tail %in% c("B","W"))
```

## Find counts & proportions, put in a table

The `janitor::tabyl()` function gets this pretty close to contingency table format...

```{r}
tail_counts <- utst %>% 
  janitor::tabyl(site, tail)

# And you can get a table with both counts & proportions! 
tail_proportions <- tail_counts %>% 
  adorn_percentages() %>% 
  janitor::adorn_pct_formatting(digits = 2) %>% 
  adorn_ns()
```

## Make contingency table and run chi-square

First, we see that the site is its own variable - when we run chi-square, we really only want the counts in the table. 

We can use `column_to_rownames` to convert entries in a column to stored rownames: 

```{r}
lizard_ct <- tail_counts %>% 
  column_to_rownames(var = "site")
```

**Now** this is in shape to run chi-square. 

What are we asking? Are site and tail breakage independent? In other words: is there a significant association between site and tail breakage? 
```{r}
lizard_x2 <- chisq.test(lizard_ct)
lizard_x2

lizard_tidy <- tidy(lizard_x2)
```
There is a significant association between site and lizard tail breakage ($\chi$^2^(`r lizard_tidy$parameter`) = `r round(lizard_tidy$statistic,2)`, *p* = `r round(lizard_tidy$p.value, 3)`). 

Remember: you'd also want to report the actual counts and proportions in each group! 

## END Part 2
